\bfseries\LARGE\singlespacing{Transmission of airborne respiratory viral infections with and without air cleaners in a Swiss school under non-pandemic conditions in 2023: A modeling study of epidemiological, environmental, and molecular data}

### 2023-07-10 doc/paper/manuscript.tex
Updates from Overleaf

\title{\bfseries\LARGE\singlespacing{Transmission of airborne respiratory infections with and without air cleaners in a Swiss school under non-pandemic conditions in 2023: A modeling study of epidemiological, environmental, and molecular data}}

@@

%TC:newcounter abst Words in abstract
%TC:envir abstract [] abst
\begin{abstract}\normalfont
\noindent\textbf{Background:} Airborne respiratory infections can be transmitted in crowded indoor settings, including schools. Infection control measures such as air ventilation or filtration can reduce the risk of airborne transmission. The aim of this study was to evaluate the effectiveness of portable HEPA-air filtration devices (air cleaners) in reducing the transmission of respiratory infections in schools using a multiple-measurement approach. \medskip

@@

The transmission of respiratory infections such as SARS-CoV-2 and influenza presents significant challenges in terms of mitigation and control. Person-to-person transmission primarily occurs through the release of respiratory particles containing the respective viruses. The recent focus has been on small respiratory particles ($\leq$ 5$\mu$m) called aerosols, which have been found to carry the majority of viruses during respiratory activities\cite{Fennelly2020}. Unlike larger respiratory droplets that tend to settle quickly, aerosols can remain suspended in the air for several hours and their respiratory plumes can reach more than 2m\cite{Coleman2022,Wang2020,Heneghan2021}, thus exceeding the recommended rules for social distancing by public health organizations\cite{Bazant2021PNAS,Trivedi2021PhyFluids,Lindsley2010CID}.

The generation, spread, and concentration of bioareosols can be influenced with infection control measures. For example, compulsory face mask wearing reduced the transmission of SARS-CoV-2 during the COVID-19 pandemic\cite{Gettings2021,Heinsohn2022,Leech2022PNAS,Mitze2020PNAS}. More generally, improved ventilation systems are considered crucial for healthy indoor environments both during pandemic and non-pandemic times\cite{Wang2021,Morawska2021}. Portable air-filtration devices (air cleaners) may present another cost-effective alternative to upgrading ventilation systems, but the effects of air cleaners on the transmission of respiratory infections are less clear. Various studies showed that air cleaners reduce the concentration of aerosols and particle matter in indoor air\cite{Park2020Build,Buising2022InfContr,Ren2021Dent,Banholzer2023PLoSMed}. These findings suggest potential protection from respiratory infection as long-term exposure to higher aerosol and particle matter mass concentration have been associated with increased risks of respiratory infections\cite{Gordon2014Resp} and, more recently, with the incidence and mortality of COVID-19\cite{Kelly2023Atmos,Veronesi2022OccEnvMed,Travaglio2021EnvPoll,DeAngelis2021EnvRes}. Furthermore, a recent study showed that air cleaners effectively removed SARS-CoV-2 bioaerosols in a hospital ward\cite{Morris2022}. As a result, studies have simulated the potential of air cleaners in reducing indoor transmission risks of SARS-CoV-2\cite{Lindsley2021} and other respiratory viruses\cite{Cortellessa2023Build}. However, empirical evidence linking the use of air cleaners with a reduced risk of infection is lacking. Thus, it remains unclear whether the reduction in particle concentrations and the removal of bioaerosols from the air translates into lower transmission of respiratory infections.

We used a multiple-measurement approach to study the transmission of respiratory viruses in school rooms under non-pandemic conditions in the winter of 2023. We collected epidemiological (absences probably related to respiratory infections), environmental (CO$_2$, aerosol and particle concentrations), and molecular data (detection of respiratory viruses in bioaerosol and human saliva samples) over a seven-week study period from January to March 2022 in two classes of a secondary school in Switzerland. We used a cross-over study design to assess the effectiveness of air cleaners as an infection control measure. For this, we estimated changes in particle concentrations, the positivity rate for respiratory viruses, and the risk of infection.

@@

>>>>

\title{\bfseries\LARGE\singlespacing{Transmission of airborne respiratory viral infections with and without air cleaners in a Swiss school under non-pandemic conditions in 2023: A modeling study of epidemiological, environmental, and molecular data}}

@@

%TC:newcounter abst Words in abstract
%TC:envir abstract [] abst
\begin{abstract}\normalfont
\noindent\textbf{Background:} Airborne respiratory infections can be transmitted in crowded indoor settings, including schools. Infection control measures such as air ventilation or filtration can reduce the risk of airborne transmission. The aim of this study was to evaluate the effectiveness of portable HEPA-air filtration devices (air cleaners) in reducing the transmission of respiratory viral infections in schools using a multiple-measurement approach. \medskip

@@

The transmission of respiratory infections such as SARS-CoV-2 and influenza presents significant challenges in terms of mitigation and control. Person-to-person transmission primarily occurs through the release of respiratory particles containing the respective viruses. The recent focus has been on small respiratory particles ($\leq$ 5$\mu$m) called aerosols, which have been found to carry the majority of viruses during respiratory activities\cite{Fennelly2020}. Unlike larger respiratory droplets, which tend to settle quickly, aerosols can remain suspended in the air for several hours and their respiratory plumes can reach more than 2 meters\cite{Coleman2022,Wang2020,Heneghan2021}, exceeding the social distancing rules recommended by public health organizations\cite{Bazant2021PNAS,Trivedi2021PhyFluids,Lindsley2010CID}.

The generation, spread, and concentration of bioareosols can be influenced by infection control measures. For example, compulsory face mask wearing reduced the transmission of SARS-CoV-2 during the COVID-19 pandemic\cite{Gettings2021,Heinsohn2022,Leech2022PNAS,Mitze2020PNAS,Banholzer2021PLOS}. In addition, improved ventilation systems are critical for a healthy indoor environment both during pandemic and non-pandemic times\cite{Wang2021,Morawska2021}. Portable HEPA-air filtration devices (air cleaners ) may present another cost-effective alternative to upgrading ventilation systems, but the effects of air cleaners on the transmission of respiratory infections are less clear. Various studies showed that air cleaners reduce the concentration of aerosols and particle matter in indoor air\cite{Park2020Build,Buising2022InfContr,Ren2021Dent,Banholzer2023PLoSMed}. These findings suggest potential protection against respiratory infection, as long-term exposure to higher aerosol and particle matter mass concentrations has been associated with increased risk of respiratory infections\cite{Gordon2014Resp} and, more recently, with the incidence and mortality of COVID-19\cite{Kelly2023Atmos,Veronesi2022OccEnvMed,Travaglio2021EnvPoll,DeAngelis2021EnvRes}. Furthermore, a recent study showed that air cleaners effectively removed SARS-CoV-2 bioaerosols in a hospital ward\cite{Morris2022}. Simulation studies have demonstrated the efficacy of air cleaners in reducing the risk of indoor transmission  of SARS-CoV-2\cite{Lindsley2021} and other respiratory viruses\cite{Cortellessa2023Build}.  We have studied the effectiveness of air cleaners in reducing the transmission of SARS-CoV-2 transmission in schools during the pandemic\cite{Banholzer2023PLoSMed}, but the analysis was limited by the study design (installation of air cleaners at the end of the Omicron wave when most students were already infected). In conclusion, empirical evidence linking the use of air cleaners with a reduced risk of infection is lacking. It remains unclear whether reducing particle concentrations and removing bioaerosols from the air will translate in reduced transmission of respiratory infections.

We used a multiple-measurement approach to study the transmission of viral respiratory viruses and the effect of aircleaners using a cross-over study design in school rooms under non-pandemic conditions in the winter of 2023. We collected epidemiological (absences probably related to respiratory infections), environmental (CO$_2$, aerosol and particle concentrations), and molecular data (detection of respiratory viruses in bioaerosol and human saliva samples) during a seven-week study period from January to March 2022 in two classes of a secondary school in Switzerland. We determined changes in particle concentrations, respiratory virus positivity rates, and modeled the risk of infection.

@@

\subsection{Comparison with previous study}

In a previous study\cite{Banholzer2023PLoSMed}, we estimated the effectiveness of mask wearing and air cleaners during the SARS-CoV-2 Omicron wave. \Cref{fig:comparison} shows a comparison of the results. The distribution of respiratory infections is different (\Cref{fig:comparison}a) because we almost exclusively detected SARS-CoV-2 in our previous study during the COVID-19 pandemic, as compared to a range of respiratory infections in this study during non-pandemic times. Fewer respiratory infections were detected in bioaerosols and on the HEPA-filters of the air cleaners in this study (\Cref{fig:comparison}b).
The reductions in particle concentrations were similar (\Cref{fig:comparison}c). The risk of infection tended to be lower with air cleaners in this study where air cleaners were analyzed using a cross-over design, but not in our previous study where air cleaners were installed only at the end of the study when most students had already been infected with SARS-Cov-2(\Cref{fig:comparison}d).

\begin{figure}[!htpb]
    \includegraphics[width=\linewidth]{../../results/epi-data/cases_by_week.pdf} 
    \includegraphics[width=\linewidth]{../../results/epi-data/avoided-infections.pdf}
    \caption{\textbf{Analysis of epidemiological data and estimated relative risk of infection with air cleaners.}. \textbf{(a)}~Weekly number of respiratory cases with and without air cleaners. \textbf{(b)}~Estimated mean number of infections (posterior mean as dot and 50\%, 80\%, and 95\%-CrI as shaded areas) if, in both classes, air cleaners had been installed throughout the study or not been installed.}
    \label{fig:comparison}
\end{figure}
